\documentclass{article}
\input{macros}
\usepackage{enumitem,amssymb}
\newlist{todolist}{itemize}{2}
\setlist[todolist]{label=$\square$}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\newcommand{\done}{\rlap{$\square$}{\raisebox{2pt}{\large\hspace{1pt}\cmark}}%
\hspace{-2.5pt}}
\newcommand{\wontfix}{\rlap{$\square$}{\large\hspace{1pt}\xmark}}

\begin{document}
\title{ITP ToDo}
\maketitle
\section*{February 2026}
\subsection*{Must be done before submission}
\begin{todolist}
    \item Keep to the page count. Note in particular the table with counterexamples in the hierarchy section as we've had trouble with text overlapping there. Clearpage was the solution but currently this is leading to a lot of unused space.
    \item [\done] Can we get rid of counterexample 9 ? This will clear a row in the counterexamples table and give us a few lines of space.
    \item [\done] Clean the section added to address a reviewer comment in the introduction. Currently it has blue text accompanying it.
    \item Get the code in a good state for submission.
    \item Write a section on our new work.
    \item Show why our cofinality defn matches the terese defn. 
    \item [\done] Theorem 1.2.3 ii: we claim the original proof is inherently classical without backing it up. We should rewrite this remark. We should say what's in the apcomment there (but tidy it up). 
    \item Try and bring space down by: rewording sentences to save lines. IN the definitions section using less line space for the definitions. Add a 2 cm space before each mathematical definition if possible so they are less left justified. 
    \item on conversaion uniqu normal form property get rid of `all'. 
    \item reduce cofinality definition down by a line or two. 
    \item hierarchy section    `the same property is similarly required in order to show that' get rid of in order to. 
    \item Definitions 27 and 28 have a fair bit of space to prune/ 
    \item cofinality equivalence: Note that the above defn is equivalent to the one in Terese. The use of reflexive closure makes the definition equivalent to the one in Terese. This is to allow for the possibility of x being normalizing and the reduction sequences from x being finite. Needed to make this equivalent to :

    ```Note the use of reflexive closure $R^r$ in the definition above. This is to allow for the possibility of a finite reduction sequence (e.g. for normalizing x), needed to make this equivalent to the definition in Terese.'''
    \item Give a very quick description of the new example of the TRS for generalized Newman's lemma applies but classical does not. \sacomment{Given in the generalized newman's lemma section.}
    \item Delete references to first order TRS.
\end{todolist}
\subsection*{Research}
\begin{todolist}
    \item [\done] Complete the TRS formalization
    \item Once TRS completed we should formalize some examples.
    \item Try to bind the formalization of the lambda calculus to the formalization of ARS. (Including the lambda calc in the LAM repo to the FPLA repo).
    \begin{itemize}
        \item Prove FB and RDec, get SN to WN?
        \item
    \end{itemize}
    \item Formalize undecidable lambda theories (at least one as an example).
\end{todolist}
\subsection*{Writing}
\begin{todolist}
    \item Complete seq minus
    \item Go through the comments from PLDI and address those that need addressing.
    \item Add high priority items from the old todo to this.
    \item [\done] (high priority) Having shown that WFmin implies WFacc when R is non empty we need to update the paper. Specifically we should update the text in the WF section and the diagram to incorporate the new implication.  \sacomment{Added to the diagram and R empty added to the table. Check definition and the text added to the paper.}
    \item (Review 3) * From the formalization, the authors could explain some features and syntax of Agda for readers unfamiliar with this proof assistant. What were the main attributes of Agda that helped you in the formalization process? In line 204, the paper brings the definition of an element being finitely branching and states that the formulation of such a definition is given in line 206. FB a seems to be a set (in line 206), not a boolean. Also, in the formalization of FB a, the authors use R b a instead of R a b. What elements of Agda allow such specification capturing exactly Def. 12.
    \apcomment{Note that the target sort of the finitely branching predicate FB is Set. The set “FB a” represents, according to the Curry-Howard isomorphism, the type of proofs that the element a is finitely branching. Such a proof is witnessed by a list of elements of A containing every R-successor of a.}
    \item [\done] Make explicit on our table in the paper that we can go from our weakest definition (wfcornotnot) to our strongest (wfmin) via classical assumptions (EM). We think we already mention that all the definitions are classically equivalent in the paper (do we?).  Also make explicit the implication from wfmin to wfacc. \sacomment{Arrow added to diagram and EM added to the table. Check EM definition. I've not added anything to the main text as we already mention that classically the definitions are equivalent.}
    \item [\done] We have shown wfmindne not not implies wfacc not not with accDNE. Now need to update the paper/graph to reflect this. \sacomment{Added to the graph and added to the remarks section. Q1: Is the `or' easy to understand on the arrow? Q2: Does it make sense to say accessibility is not not closed `on' the relation or should it be described differently? }
    \item Does EM imply all of our classical hypothesis? (We've shown it does for some of them)
    \item The question of whether SN implies WN  in a first order TRS relates to the question of whether we can show isMinDec holds. If isMinDec holds then SN implies WN. (we've shown this holds in implications.agda)
    \item We should try and find an ARS example where generalized NL holds but not the standard definition. This will help illustrate why it is useful! \sacomment{CE:8 in our paper covers this (two objects with relations to one another). Is this too much of a toy example?}
    \item [\done] * L5: The abstract talks about "Abstract Reduction Systems", but the introduction mentions "Abstract Rewriting Systems". Make it consistent through the paper! We use rewriting.
    \item [\done] I recommend against citations in the abstract, mostly because they will be shown in contexts where the bibliography is not readily available, i.e., in summarization services or on the download page of the article. (Need to remove the abstract citations)
    \item [\done] * L7: "their relationships between each other" reads awkwardly as a phrase. (rewrite)
    \item [\done] * L18-18: "as presented in Term Rewriting Systems [6]" it seems a bit odd to refer to the book by name like this. Since the reference is so foundational to the work, perhaps you can expand it to say "as presented in the book by Bezem, Klop and De Vrijer" or something to that effect. 
    \item [\done] * L38: I recommend preceding~\cite with ~ to prevent a citation from appearing as the very first thing on a new line.  (should be applied throughout the paper)
    \item [\done] * L93: Maybe this paragraph, being just one sentence, can be merged with the one that precedes it. (Sentence/para starting "As long as the code does not use")
    \item [\done] WN and SN are introduced as abbreviations before the terms are defined. 
    \item [\done] "Completeness" is introduced without definition: "   The result is a broader ontology of ARS concepts, revealing interrelationships between conditions for completeness."
    \item [\done] * L237, L239: Why are the parentheses dropped in "sk"? I guess this is meant to read as "s applied to k", but elsewhere the parentheses do appear. \sacomment{Added the parens.}
    \item [\done] * L281: I am still a bit surprised to see that, in this Agda formalization of FB, the relation R seems to work "the other way around" -- i.e., I would have expected "R a b" rather than "R b a". (We should make a clarifying Agda remark at this point?) 
    \item [\done] "... and further discussion of [\texttt{isWFseq-}] are provided ..." \sacomment{is it just the use of `are' rather than `is' here that they're highlighting? If so, I've changed it.}
    \item [\done] * L361: The text spills into the margin here. (We should go through the whole paper to ensure this doesn't occur elsewhere too)
    \item * L372: "minimalizing" is an awkward term; how about "minimizing"? \sacomment{Shall we apply this change through the paper?}
    \item * L449-450: These footnotes seem to apply to Table 1 already, why not specify them there? (This worked better when the tables were on one page together. If we can't get them to fit on a table together it does make sense to replicate the footnotes.)
    \item [\done] * L454: The "counterexamples" subsection is somewhat odd. Why would you put its contents in a separate subsection if there is no accompanying prose? (I think this is a good point but I also think it is useful to highlight the CE section. Solutions: We could make a sub sub section? Or write an explanatory paragraph/caption)
    \item Reviewer has highlighted that this is a symbol soup that could probably be expressed in prose: "(There is no loss of generality in this; any global well-foundedness property...)" (end of this line is the particular culprit) \sacomment{I'm not sure how to express this section in english rather than via the symbols. But I agree that it could be made clearer!}
    \item We could address the why and how more. In explaining sec 5 we could see whether the bullet point list could be expanded into a theorem with a bunch of summaries of the results in the table. In the proof we can focus on the more interesting and `leave the more trivial to the agda code.'
    \item I've added the definition of accessibility to the section where we define strongly normalizing as per a reviewer comment. I've reversed the relation in this context. Should we address the fact that the relation is in different directions in the two places we define accessibility?
    \item * L207, L210: The symbol `triple equals' on this line is undefined. We can make clear in the paper what we are referring to with `triple equals': The relation  in the
    last two items refers to intensional equality defined inductively as the least reflexive relation, also known as the Martin-Löf identity type.
    \item For the cofinality property we need to make clear why our definition matches the one in Terese. Ideally we show equivalence. 
    \item [\done] remove the redundant discussion about the straightforward propositions.
    \item We mention that theorem 1-2-3 ii is `inherently classical'. We should make clear why we think this by displaying a constructive taboo.
    \item A number of reviewers are specifically interested in how our paper differs from the Color paper (trs formalized in coq/roq). From reading that paper some more I think we could characterise their focus as being on creating a formalization of certain key results in term rewriting so as to act as a form of verification of termination certificates. It does not do what our paper does: look into the taxonomy of the important properties of term rewriting and explore new relations between them. Maybe a lazy summary but: They make available known results of term rewriting through their formalization. We use our formalization to explore new results in rewriting. 
    \item A few of the reviewer comments mentioned that we should focus more on decidable relations. We should either accept decidability or make clear why we don't want to assume decidability.
    \item See the first reviewer comment about the two motivating examples we gave. For them, these examples were insufficient. We either need to better present them or change them.
    \item We are asked what our well foundedness figure would look like if we assumed decidability of R and being a normal form is decidable. Do we want to address this?
    \item We are asked by a reviewer whether our Agda formalization allows for extractable of executable algorithms. We need to make this clear, preferably through a simple example.
    \item Reviewer comment: we state we avoid using axioms/postulates that are non constructive. They don't think they're used in the first place. Can we find examples where they are used?
    \item We are asked by a reviewer to be explicit about whether we are formalizing anything that hasn't already been formalized.
    \item For readers who don't know Agda, we should provide some sort of explanation of the proofs for our new results (generalized newman's lemma, our constructive variations of existing proofs).
    \item Make the paper less of a readme of the code and go into the why and how of things more. In explaining sec 5 we could see whether the bullet point list could be expanded into a theorem with a bunch of summaries of the results in the table. In the proof we can focus on the more interesting and `leave the more trivial to the agda code.'
    \item  MF and WN iff NF (see in hierarchy section) should be an intersection rather than logical and (change to be made in paper and agda code). Similarly it should be the double arrow iff. Check through paper for similar changes.  \sacomment{I've made the change to the paper but haven't changed the agda code as I'm unable to get the type right. Should the remarks on the hierarchy table (remark 24) also make use of set notation rather than logic notation?}
    
\subsection*{Removed writing}
I'm going to remove parts of the paper to make space for the new section. I'll leave them in the original place but commented out. Here I will list what was removed to hopefully give a quick overview of the changes
\begin{enumerate}
    \item (Sec:Definitions) ``The file \texttt{ClosureOperators.agda} verifies key properties of closure operators on relations, including their monotonicity, preservation of equivalence, and interactions such as inclusions and equivalences between different closures."
    \item (Sec:Definitions) `` The definitions in this subsection are key to formalizing Theorem 1.2.3 in \terese.''
    \item (Sec:Definitions) ``$\FB$ plays an important role in our discussion of well-foundedness, see Section \ref{sec:Well-foundedness}.''
    \item (sec:hierarchies) ``Trivially, the following holds.''
    \item (sec:WellFounded) ``(There is no loss of generality in this; any global well-foundedness property $gWF(A;R)$  restricts to a predicate $WF_R$ on $A$ via $WF_R(x) = gWF(\Sigma y {:} A. R^* x y;R')$, where $R'((y,\rho),(z,\sigma)) = R(y,z)$; now $gWF(A;R) \leftrightarrow \forall x. WF_R(x)$.)``. (I think this is useful to add but not necessary) 
    \item (sec:wellfounded) ``Recall the definition of an $R$-decreasing sequence from Section~\ref{sec:Definitions}:this is a function $s : \nat \to A$ satisfying $R(s(k+1))(s(k)) \in R$ for all $k \in \nat$.''
    \item (seq:wellfounded) ``In terese, the $WFseqm$ definition is used, as in the classical mathematics textbooks. Constructively, this is not a useful definition, as it does not provide any positive computational content and requires an explicit sequence to be defined.  This definition appears to be strictly weaker than $WFseqnn$, asserting that every sequence not-not contains a non-decreasing index.''
    \item (sec:Wellfounded) ``We have also considered the statement that the complement of every coreductive predicate
  has a least element.  This statement and its $\lnot\lnot$-version are both equivalent to $\WFcorm$
  (see \texttt{WFCoreductiveImplications.agda}), so we have not included them in the figure. We also did not include $WFind$ from Definition \ref{def:WFproperties}. Logically, it is trivially equivalent to the accessible one, but its Agda formalization has a defect: the inductive well-foundedness property lands in the larger universe $\mathtt{Set}_1$, due to  quantification over all predicates on $A$.  This defect is also present in $WFmin$, $WFminDNE$, and $WFcor$.''
 \end{enumerate}


\end{todolist}


\end{document}
