\makeatletter
\def\input@path{{./tex/acmart/}}
\makeatother


\documentclass[
  acmsmall,
  screen,
  review,      % remove this for camera-ready
  anonymous,   % remove for camera-ready
  nonacm       % PACMPL papers use PACMPL style but are technically non-ACM proceedings submissions
]{acmart} 

% --- metadata ---
\title{A Formalization of Abstract Rewriting}

\author{Samuel Arkle}
\affiliation{
  \institution{Appalachian State University}
  \country{USA}
}
\email{arklesd@appstate.edu}

\author{Andrew Polonsky}
\affiliation{
  \institution{Appalachian State University}
  \country{USA}
}
\email{polonskya@appstate.edu}

% --- Keywords, CCS ---
\keywords{Abstract rewriting, Formalization, Confluence, Strong normalization, Newman's Lemma, Well-foundedness}

% \begin{CCSXML}
% <ccs2012>
%    <concept>
%       <concept_id>10003752.10003790.10011741</concept_id>
%       <concept_desc>Theory of computation~Type theory</concept_desc>
%       <concept_significance>300</concept_significance>
%    </concept>
% </ccs2012>
% \end{CCSXML}

\ccsdesc[300]{Theory of computation~Type theory}

% --- packages & macros ---
\input{macros} 
% \newtheorem{definition}{Definition}
% \newtheorem{notation}{Notation}
\newtheorem{remark}{Remark}
\theoremstyle{acmdefinition}
\newtheorem{notation}{Notation} 
% \newtheorem{notation}[theorem]{Notation} % Can't get this to work

% \supplement{https://github.com/DrPolonsky/FPLA/tree/main} %TODO: Can't get this reference to the code to work.
\usepackage{enumitem}
% --- begin document ---
\begin{document}

\begin{abstract}
    We present a constructive formalization of Abstract Reduction Systems (ARS)
    in the Agda proof assistant, focusing on the results given in the standard
reference on Term Rewriting Systems \cite{Terese}. We define a taxonomy of
concepts related to termination and confluence and investigate their
relationships between each other and their classical counterparts. We identify,
and eliminate where possible, the use of classical logic in the proofs of
standard ARS results. Our analysis leads to refinements and mild
generalizations of classical termination and confluence criteria. Finally, we
investigate logical relationships between several notions of termination,
arising from different formulations of the concept of a well-founded relation.
\end{abstract}

\maketitle

\input{Introduction.tex}
\input{Definitions.tex}
\input{Formalization-pldi.tex}
\input{Implications.tex}
\input{Well-foundedness.tex}
\input{Conclusion.tex}

% \bibliographystyle{ACM-Reference-Format}
% \bibliography{references} % create references.bib instead of inline bib



\begin{thebibliography}{10}
    \bibitem{Terese}
    Marc Bezem, Jan Willem Klop, and Roel de Vrijer.
    \newblock {Term Rewriting Systems}.
    \newblock {Cambridge University Press, 2003}

    \bibitem{HoTT}
    Peter Aczel, Benedikt Ahrens, Thorsten Altenkirch, Steve Awodey,
    Bruno Barras, Andrej Bauer, Yves Bertot, Marc Bezem, Thierry Coquand,
    Eric Finster, et al.
    \newblock{Homotopy Type Theory: Univalent Foundations of Mathematics}.
    \newblock{The Univalent Foundations Program, Institute for Advanced Study, 2013}
    
    \bibitem{QIIT}
    Thorsten Altenkirch, Paolo Capriotti, Gabe Dijkstra, Nicolai Kraus,
    and Fredrik Nordvall Forsberg.
    \newblock{Quotient inductive-inductive types}.
    \newblock{Proceedings of the International Conference on Foundations of Software Science and Computation Structures},
    \newblock{pages 293--310. Springer International Publishing, Cham, 2018}
    
    \bibitem{CTT}
    Thierry Coquand, Simon Huber, and Anders Mörtberg.
    \newblock{On higher inductive types in cubical type theory}.
    \newblock{Proceedings of the 33rd Annual ACM/IEEE Symposium on Logic in Computer Science},
    \newblock{pages 255--264, 2018}
    
    \bibitem{mimram}
    Mimram, Samuel
    \newblock{Program= Proof}.
    \newblock{2020}.
    \newblock{isbn = 9798615591839}.
    \newblock{Independently Published}
    
    \bibitem{Bezem}
    Marc Bezem, Thierry Coquand, Keiko Nakata, and Erik Parmann.
    \newblock{Realizability at Work: Separating Two Constructive Notions of Finiteness}.
    \newblock {In 22nd International Conference on Types for Proofs and Programs (TYPES 2016)}.
    \newblock{Leibniz International Proceedings in Informatics (LIPIcs), Volume 97, pp. 6:1-6:23, Schloss Dagstuhl – Leibniz-Zentrum für Informatik (2018)}.
    \newblock{https://doi.org/10.4230/LIPIcs.TYPES.2016.6}
    
    \bibitem{Berardi}
    Stefano Berardi
    \newblock{Classical and Intuitionistic Arithmetic with Higher Order Comprehension Coincide on Inductive Well-Foundedness}.
    \newblock{In 24th EACSL Annual Conference on Computer Science Logic (CSL 2015) }.
    \newblock{Leibniz International Proceedings in Informatics (LIPIcs), Volume 41, pp. 343-358}.
    \newblock{Schloss Dagstuhl – Leibniz-Zentrum für Informatik (2015)}.
    \newblock{https://doi.org/10.4230/LIPIcs.CSL.2015.343}
    
    \bibitem{AgdaLibraryList}
    Agda Standard Library: List
    \newblock{https://agda.github.io/agda-stdlib/v2.2/Agda.Builtin.List.html}
    \newblock{Accessed: 8th July 2025}
    
    \bibitem{AgdaLibraryRewriting}
    Agda Standard Library: Rewriting
    \newblock{https://agda.github.io/agda-stdlib/v1.7.3/Relation.Binary.Rewriting.html}
    \newblock{Accessed: 5th November 2025}
    
    \bibitem{isabelleFormalization}
    Christian Sternagel and René Thiemann
    \newblock{https://www.isa-afp.org/entries/Abstract-Rewriting.html}
    \newblock{Accessed: 5th November 2025}
    
    \bibitem{trs}
    André Galdino, Mauricio Ayala-Rincón
    \newblock{http://trs.cic.unb.br/}
    \newblock{Accessed: 5th November 2025}
    
    \bibitem{coqARS}
    
    Blanqui, Frédéric, and Adam Koprowski.
    \newblock{“CoLoR: A Coq Library on Well-Founded Rewrite Relations and Its Application to the Automated Verification of Termination Certificates.”}
    \newblock{Mathematical Structures in Computer Science 21, no. 4 (2011): 827–59.}\\
    \newblock{\url{https://doi.org/10.1017/S0960129511000120}}
    
    \bibitem{AITC}
    \newblock{Annual International Termination Competition}
    \url{https://termination-portal.org/wiki/Termination_Competition}

  \end{thebibliography}
  
\end{document}
% Plan of the paper:

% 1. Introduction: 2 pages.
% 2. Definitions/setup: 2 pages. [WIP]
% 3. Formalization of Chapter 1 Terese, including Newman's Lemma: 2 pages [Still needs to be done]
% - Statements (not proofs!) of key propositions and theorems in Agda code
% 4. Implications, hierachies, figures, and counterexamples: 4-5 [draft in report]
% 5. Classical principles and well-foundedness: 3 pages
% 6. Miscellaneous facts: 1 page
% 7. Related and future work: 1 page
